RNeXML: a package for reading and writing richly annotated phylogenetic, character, and trait data in R

### 2015-01-03 manuscripts/manuscript.Rmd
Rewrite of the publishing to repositories section.

Data archiving is increasingly required by journals in evolutionary biology and 
biodiversity (e.g. @Rausher_2010).  The burden of preparing and submitting properly
annotated data to archives continues to be a significant barrier [@Tenopir_2011; @Stodden_2014],
and many phylogenetic trees upon which studies are based are inaccessible or lost 
to the research community [@Drew_2013].  `RNeXML` seeks to lower these barriers
by providing functions to immediately archive NeXML files in scientific repositories.

@@

Robust data archiving need not involve immediate release of the data. The `figshare` 
repository supports the option of `visibility = "private"`, allowing secure 
publication that provides a way to backup and share data with collaborators 
prior to public release.

# Software reuse and support

The `RNeXML` package is available through the CRAN repository under
a CC0 license. At the time of writing, the current version is `r packageDescription("RNeXML", fields="Version")`. 
The current development
branch can be found on Github at [https://github.com/ropensci/RNeXML]().
See the NEWS file for changes in recent versions. Users are encouraged
to submit bug reports or feature requests in the issues log.  Contact
`ropensci-discuss@googlegroups.com` or the package maintainer for help,
or email the phylogenetics R users group list at `r-sig-phylo@r-project.org`.

There are many promising directions for further development of the
`RNeXML` package.  The project's [Milestones page](https://github.com/ropensci/RNeXML/issues/milestones) page provides a list of
steps currently planned or under consideration.  In particular, these
milestones include increased support for interactive mode of entering data
(wizards), conversions to and from other existing R phylogenetics formats
such those found in the `phylobase` package [@phylobase],
and support for integration of semantics defined through the
Web Ontology language (OWL), as discussed in the Challenges section.

## Future Directions

For convenience, the RNeXML package can parse compressed XML binaries (gzipped XML) directly, facilitating work with very large data files.  As most R functions, especially those in common phylogenetics packages, require in-memory data structures, RNeXML currently loads this data directly into memory. Future work could facilitate manipulation of NeXML data directly from the disk.

>>>>

Data archiving is increasingly required by scientific journals,
including in evolutionary biology, ecology, and biodiversity
(e.g. @Rausher_2010). The effort involved with preparing and
submitting properly annotated data to archives remains a notable
barrier to the broad adoption of data archiving and sharing as a
normal part of the scholarly publication workflow
[@Tenopir_2011; @Stodden_2014]. In particular, the majority of
phylogenetic trees published in the scholarly record are inaccessible
or lost to the research community [@Drew_2013].

One of `RNeXML`'s aims is to promote the archival of well-documented
phylogenetic data in scientific data repositories, in the form of
NeXML files. To this end, the method `nexml_publish()` provides an API
directly from within R that allows data archival to become a step programmed
into data management scripts, rather than a separate manually executed
act. Initially, the method supports the data repository Figshare (http://figshare.com):

@@

Figshare also supports "private" visibility, allowing to securely backup
data to a repository and to share them with collaborators prior to public release.

# Future Directions
